---
title: "Smart_homes"
author: "Stacey Jackson"
date: "16/04/2019"
output: word_document
---
####1. INSTALL PACKAGES
```{r}

library(RMySQL)
library(dplyr)


library(lubridate)

library(ggplot2)

library(tidyr)
```



```{r}
con = dbConnect(MySQL(), user='deepAnalytics', password='Sqltask1234!', dbname='dataanalytics2018', host='data-analytics-2018.cbrosir2cswx.us-east-1.rds.amazonaws.com')
```

Upload each year and list attributes in each table
```{r}
yr_2006 <- dbGetQuery(con, "SELECT * FROM yr_2006")
dbListFields(con,"yr_2006")


```
####2. DATA FOR EACH SUB-METER
####2.1 Download date, time and the sub-metering attributes for each sub-meter
```{r}
yr_2006SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2006")
yr_2007SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2007")
yr_2008SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2008")
yr_2009SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2009")
yr_2010SELECT <- dbGetQuery(con, "SELECT Date, Time, Global_active_power, Sub_metering_1, Sub_metering_2, Sub_metering_3 FROM yr_2010")
```
####3. INVESTIGATE THE DATA
####3.1 2006
```{r}
str(yr_2006SELECT)
summary(yr_2006SELECT)
head(yr_2006SELECT)
tail(yr_2006SELECT)
```
####3.2 2007
```{r}
str(yr_2007SELECT)
summary(yr_2007SELECT)
head(yr_2007SELECT)
tail(yr_2007SELECT)
```
####3.3 2008
```{r}
str(yr_2008SELECT)
summary(yr_2008SELECT)
head(yr_2008SELECT)
tail(yr_2008SELECT)
```
####3.4 2009
```{r}
str(yr_2009SELECT)
summary(yr_2009SELECT)
head(yr_2009SELECT)
tail(yr_2009SELECT)
```
####3.5 2010
```{r}
str(yr_2010SELECT)
summary(yr_2010SELECT)
head(yr_2010SELECT)
tail(yr_2010SELECT)
```
####4. CREATE ONE DATAFRAME - ONLY YEARS THAT ARE COMPLETE
```{r}
newDF <- bind_rows(yr_2007SELECT, yr_2008SELECT, yr_2009SELECT)
```
####4.1 Investigate the new dataframe and rename columns
```{r}

names(newDF)[4] <- "Kitchen"
names(newDF)[5] <- "Laundry_Room"
names(newDF)[6] <- "Heating_Water"

str(newDF)
summary(newDF)
head(newDF)
tail(newDF)
```
####4.2 Change Sub-meters' readings to kW/hr
```{r}
newDF$Kitchen <- newDF$Kitchen/1000 
newDF$Laundry_Room <- newDF$Laundry_Room/1000 
newDF$Heating_Water <- newDF$Heating_Water/1000 

head(newDF)

```
####4.3 Create Rest_of_House to cover the rest of the house
```{r}
newDF$Rest_of_House<-newDF$Global_active_power/60-(newDF$Kitchen+newDF$Laundry_Room+newDF$Heating_Water)

head(newDF)
```

####5.TIME
####5.1 Combine Date & Time attribute values in a new attribute column
```{r}
newDF <-cbind(newDF,paste(newDF$Date,newDF$Time), stringsAsFactors=FALSE)
```
####5.2 Give the new attribute in the 8th column a header name 
```{r}
colnames(newDF)[8] <-"DateTime"
```
####5.3 Move the DateTime attribute within the dataset
```{r}
newDF <- newDF[,c(ncol(newDF), 1:(ncol(newDF)-1))]
head(newDF)
tail(newDF)
```
####5.4 POSIXct conversion
```{r}
## Convert DateTime from POSIXlt to POSIXct 
newDF$DateTime <- ymd_hms(newDF$DateTime)
newDF$Time<-hms(newDF$Time)
newDF$Date<-ymd(newDF$Date)

## Add the time zone
#attr(newDF$DateTime, "tzone") <- "Europe/Paris"

## Inspect the data types
str(newDF)
```
####6.YEAR 
```{r}
## Create "year" attribute with lubridate

newDF$year <- year(newDF$DateTime)
```
Review your primary data frame. Was the year attribute created? What values are found in the year attribute? 
```{r}
str(newDF)
summary(newDF)
```
####6.1 Attributes for quarter, month, week, weekday, day, hour and minute
```{r}
newDF$quarter <- quarter(newDF$DateTime)
newDF$month <- month(newDF$DateTime)
newDF$week <- week(newDF$DateTime)
newDF$weekdays <- weekdays(newDF$DateTime)
newDF$day <- day(newDF$DateTime)
newDF$hour <- hour(newDF$DateTime)
newDF$minute <- minute(newDF$DateTime)

str(newDF)
summary(newDF)
head(newDF)
tail(newDF)
```
#### 7. LOOK FOR MISSING DATES
```{r}

newDF$Date2 <- as.Date(newDF$Date)
date_range <-seq.Date(min(newDF$Date2), max(newDF$Date2), by="day")
date_range[!date_range %in% newDF$Date2] 


```


####8. SUMMARY STATISTICS
####totals for each
```{r}
summary(newDF)

sum(newDF$Kitchen)
sum(newDF$Laundry_Room)
sum(newDF$Heating_Water)
sum(newDF$Rest_of_House)



```
####8.1 sub-meter1
Minimum = 0
Maxiumum = 0.082000
Mean = 0.001159
Median = 0
Standard deviation = 0.006288272
1st quartile = 0
3rd quartile = 0
```{r}
sd(newDF$Kitchen)


plot(newDF$Kitchen)
boxplot(newDF$Kitchen)

p <- ggplot(data=newDF, mapping=aes(x=Kitchen))
p + geom_histogram(bins=8)

p <- ggplot(data=newDF, mapping=aes(x=quarter, y=Kitchen))
p + geom_bar(stat="identity")

p <- ggplot(data = newDF, mapping = aes(x = month, y = Kitchen, color=Kitchen))
p + geom_smooth()

p <-  ggplot(data=newDF, mapping= aes(x = quarter, y=Kitchen, colour=year)) 
p + geom_bar(stat="identity")

p <- ggplot(data = newDF, mapping = aes(x = weekdays, y = Kitchen, fill=Kitchen))
p + geom_bar(stat="identity")
```
####8.2 sub-meter2
Minimum = 0
Maxiumum = 0.078000
Median = 0
Mean = 0.001343
Standard deviation = 0.005972199
1st quartile = 0
3rd quartile = 0.001000

```{r}
sd(newDF$Laundry_Room)
plot(newDF$Laundry_Room)
boxplot(newDF$Laundry_Room)

p <- ggplot(data=newDF, mapping=aes(x=Laundry_Room))
p + geom_histogram(bins=8)

p <- ggplot(data=newDF, mapping=aes(x=quarter, y=Laundry_Room))
p + geom_bar(stat="identity")

p <- ggplot(data = newDF, mapping = aes(x = month, y = Laundry_Room, color=Laundry_Room))
p + geom_smooth()

p <-  ggplot(data=newDF, mapping= aes(x = quarter, y=Laundry_Room, colour=year)) 
p + geom_bar(stat="identity")

p <-  ggplot(data=newDF, mapping= aes(x = year, y=Laundry_Room, colour=year)) 
p + geom_bar(stat="identity")
```
####8.3 sub-meter3
Minimum = 0
Maxiumum = 0.031000
Median = 0.001000
Mean = 0.006216
Standard deviation = 0.008341281
1st quartile = 0
3rd quartile = 0.017000

```{r}
sd(newDF$Heating_Water)
plot(newDF$Heating_Water)
boxplot(newDF$Heating_Water)

p <- ggplot(data=newDF, mapping=aes(x=Heating_Water))
p + geom_histogram(bins=8)

p <- ggplot(data=newDF, mapping=aes(x=quarter, y=Heating_Water))
p + geom_bar(stat="identity")

p <- ggplot(data = newDF, mapping = aes(x = month, y = Heating_Water, color=Heating_Water))
p + geom_smooth()

p <-  ggplot(data=newDF, mapping= aes(x = quarter, y=Heating_Water, colour=year)) 
p + geom_bar(stat="identity") 
  
```
####8.4 sub-meter4
Minimum = 
Maxiumum = 
Median = 
Mean = 
Standard deviation = 
1st quartile = 
3rd quartile = 

```{r}
sd(newDF$Rest_of_House)
plot(newDF$Rest_of_House)
boxplot(newDF$Rest_of_House)

p <- ggplot(data=newDF, mapping=aes(x=Rest_of_House))
p + geom_histogram(bins=8)

p <- ggplot(data=newDF, mapping=aes(x=quarter, y=Rest_of_House))
p + geom_bar(stat="identity")

p <- ggplot(data = newDF, mapping = aes(x = month, y = Rest_of_House, color=Rest_of_House))
p + geom_smooth()

p <-  ggplot(data=newDF, mapping= aes(x = quarter, y=Rest_of_House, colour=year)) 
p + geom_bar(stat="identity")
```

####8.5 All together
```{r}



##By year
newDF %>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(mapping= aes(x = year, y=value, colour=key)) +
  geom_bar(stat="identity") +
  facet_wrap(~key,ncol=1)

##By quarter
newDF %>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(mapping= aes(x = quarter, y=value, fill=key)) +
  geom_bar(stat="identity") +
  facet_wrap(~key,ncol=1)

##By month
newDF%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = month, y = value, color=key))+
  geom_smooth() +
  facet_wrap(~key,ncol=1)

## By day of the week
newDF%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = weekdays, y = value, color=key))+
  geom_bar(stat="identity") +
  facet_wrap(~key,ncol=1)

## By hour of day week
newDF%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = hour, y = value, color=key))+
  geom_smooth() +
  facet_wrap(~key,ncol=1)

newDF%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = hour, y = value, color=key))+
  geom_smooth() +
  coord_flip() + 
  facet_grid(~key)

newDF%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = hour, y = value, color=key))+
  geom_bar(stat="identity") +
  coord_flip() + 
  facet_grid(~key)

newDF%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = weekdays, y = value, color=key))+
  geom_bar(stat="identity") +
  coord_flip() + 
  facet_grid(~key)


```
####9. PRICES
```{r}
newDF$rate<-ifelse (newDF$Time >=hms("06:31:00") & newDF$Time<=hms("22:29:00"), 0.00263, 0.0021)

newDF$price<-newDF$rate*newDF$Global_active_power


#15 Sept 2009
newDF %>% 
  filter(Date==as.Date("2009-09-15")) %>% 
  summarise(sum(price))

#Total Jan-Aug 2009
newDF %>% 
  filter(between(Date, as.Date("2009-01-01"), as.Date("2009-08-31")))%>%
  summarise(sum(price))

#Total 1-14 Sept 2009
newDF %>% 
  filter(between(Date, as.Date("2009-09-01"), as.Date("2009-09-14")))%>%
  summarise(sum(price))

```
####10.Plots for mock-up
```{r}
#Water/heating 15 September 2009
newDF%>%
  filter(year==2009)%>%
  filter(month==9)%>%
  filter(day==15)%>%
  gather(key,value, Heating_Water) %>%
  ggplot(aes(x = hour, y = value,fill=key))+
  geom_bar(stat="identity")
  

#Kitchen September 2009
newDF%>%
  filter(year==2009)%>%
  filter(month==9)%>%
  filter(day<15)%>%
  gather(key,value, Heating_Water) %>%
  ggplot(aes(x = day, y = value, color=key))+
  geom_bar(stat="identity") 

#Kitchen Up to Sept 2009
newDF%>%
  filter(year==2009)%>%
  filter(month<9)%>%
  gather(key,value, Heating_Water) %>%
  ggplot(aes(x = month, y = value, color=key))+
  geom_bar(stat="identity") 

#Overview Up to Sept 2009
newDF%>%
  filter(year==2009)%>%
  filter(month<9)%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = month, y = value, color=key))+
  geom_bar(stat="identity")

#Overview 15 September 2009
newDF%>%
  filter(year==2009)%>%
  filter(month==9)%>%
  filter(day<15)%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = day, y = value, color=key))+
  geom_bar(stat="identity") 


#Pie chart 1-14 Sept 2009
newDF%>%
  filter(year==2009)%>%
  filter(month==9)%>%
  filter(day<15)%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = "", y = value, fill=key))+
  geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)

#Pie chart Jan-Aug 2009
newDF%>%
  filter(between(Date, as.Date("2009-01-01"), as.Date("2009-08-31")))%>%
  summarise(Kitchen=sum(Kitchen), Laundry_Room=sum(Laundry_Room), Heating_Water=sum(Heating_Water), Rest_of_House=sum(Rest_of_House))%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = "", y = value, fill=key))+
  geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)

#Pie chart 15 September 2009
newDF%>%
  filter(year==2009)%>%
  filter(month==9)%>%
  filter(day<15)%>%
  summarise(Kitchen=sum(Kitchen), Laundry_Room=sum(Laundry_Room), Heating_Water=sum(Heating_Water), Rest_of_House=sum(Rest_of_House))%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = "", y = value, fill=key))+
  geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)

#Pie chart 15 September 2009
newDF%>%
  filter(year==2009)%>%
  filter(month==9)%>%
  filter(day==15)%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = "", y = value, fill=key))+
  geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)

#Overview 15 September 2009
newDF%>%
  filter(year==2009)%>%
  filter(month==9)%>%
  filter(day==15)%>%
  gather(key,value, Kitchen, Laundry_Room, Heating_Water, Rest_of_House) %>%
  ggplot(aes(x = hour, y = value, fill=key))+
  geom_bar(stat="identity") 


```

